Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-11-13 manuscript/review_response_R1.Rmd

Lines 132-135: Wasn’t initially clear that these were manually collected data instead of another set of sensors. Maybe mention that they were collected by handheld probe at the beginning of the description instead of at the very end.

\response{Done.}

Lines 142-147: Description of standardization procedure is confusing. I had to read it three times before the procedure was clear to me. Consider rewording for clarity.

\response{Upon re-reading, we completely agree! We have substantially rewritten this section for clarity and have added some mathematical notation to make our description precise (lines xxx-xxx). Hopefully it is less confusing now.}

@@

\response{We have revised this section to make this more clear (lines xxx-xxx).}

Lines 168-174: I like the idea of using the two separate estimates of greenness and then using the one that predicts best every year. Very thorough.

\response{Thank you.}

Line 188: later, in figure 2, the authors refer to mean volumetric water content as ``VWC'' also. It would be best to use a different abbreviation for the two separate uses of volumetric water content: mean and cumulative.

@@

\response{Thank you.}

Lines 289-299: strong discussion and support for the sizes of the treatment applied. Good work.

\response{Thanks for the kind words.}

Lines 320-325: seems like authors are indicating that ecosystem response was driven by the response of the dominant species. Might be good to bring in some other work that has found similar results and discuss it a bit, for example: ``Drivers of Variation in Aboveground Net Primary Productivity and Plant Community Composition Differ Across a Broad Precipitation Gradient'' (La Pierre et al., 2016).

@@

Table 1: These are the beta coefficents from the model correct? Might be good to elaborate a bit more in the table description.

\response{Yes, these are the beta coefficients. We have added detail to the table description.}

Figure 2: Here is the use of VWC as “mean soil volumetric water content” I referred to earlier. Like I said it would probably be best to have different abbreviations for different metrics of soil moisture. Also interesting, though not the point of this paper, after three consecutive dry years with progressively decreasing ANPP it looks like the ecosystem was primed for recovery as soon as an average precipitation year hit in 2015. No evidence of drought legacy.

\response{Sorry about this mistake. We use \emph{cumulative} March-June volumetric water content for the statistical model, but in this figure we plot the estimated daily values as a time series. We have updated the figure axis label and the caption to clarify.}

@@

Figures general: differences between colors used to indicate treatments are not at all distinguishable if viewed in greyscale. Authors should consider using a color scheme with more contrast.

\response{We have increased the contrast between colors so that they are easier to distinguish if printed in black and white. They are not perfect, but they are from a color-blind friendly palette and we anticipate most readers of \emph{PeerJ} will read the paper online.}

### 2017-11-13 manuscript/usses_water_ms.Rmd
more revisions and updates fig 2 VWC label

We first averaged the observed soil moisture for each day (*d*) and plot (*i*), $x_{i,d}$.
\new{Then we standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_G x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.}
We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group.
These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.

We then modeled the daily deviation from ambient conditions of the drought and irrigation treatments using a linear mixed effects model with independent variables for treatment, season (winter, spring, summer, fall), rainfall, and all two-way interactions. 
Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
The day immediately following rainfall was also classified as rainy.
We fit the model using the ```lme4::lmer()``` function [@Bates2015] in R [@R2016], with random effects for plot group and date.
We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day.
We then used the model to predict the average daily soil moisture in the treated plots based on the average daily ambient soil moisture.
We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
See @Kleinhesselink2017b for complete details on our approach to estimating soil moisture.

@@

\begin{description}
\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
\item [Appendix 2.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
\item [Appendix 3.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
\item [Appendix 4.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
\end{description}

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=7in]{../figures/data_panels.png}
  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis. (B) Mean soil volumetric water content (VWC) in each of the three treatments during the course of the experiment. (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment. Colors in panels B and C identify the treatment, as specified in the legend of panel C.}
\end{figure}

>>>>

\new{We first averaged the observed soil moisture for each day ($d$) and plot ($i$), $x_{i,d}$.
Plots were assigned to plot groups consisting of a control, a drought, and an irrigation plot.
We standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_n x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.
These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.
We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group: $\Delta x^*_{\text{trt.}} = x^*_{\text{control}} - x^*_{\text{trt.}}$.}

We then modeled the daily deviation \new{($\Delta x^*_\text{trt.}$)} from ambient conditions of the drought and irrigation treatments using a linear mixed effects model with independent variables for treatment, season (winter, spring, summer, fall), rainfall, and all two-way interactions. 
Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
The day immediately following rainfall was also classified as rainy.
We fit the model using the ```lme4::lmer()``` function [@Bates2015] in R [@R2016], with random effects for plot group and date.
We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day.
We then estimated average daily soil moisture in the treated plots based on the average daily ambient soil moisture \new{and the model predicted daily deviation for each treatment ($\Delta x^*_\text{trt.}$)}.
We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
See @Kleinhesselink2017b for complete details on our approach to estimating soil moisture.

@@

\begin{description}
\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
\item [Appendix 2.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
\item [Appendix 3.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
\item [Appendix 4.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
\end{description}

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=7in]{../figures/data_panels.png}
  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis. (B) \new{Estimated daily} soil volumetric water content (VWC) in each of the three treatments during the course of the experiment. (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment. Colors in panels B and C identify the treatment, as specified in the legend of panel C.}
\end{figure}

### 2017-11-13 manuscript/usses_water_ms.tex
more revisions and updates fig 2 VWC label

We first averaged the observed soil moisture for each day (\emph{d}) and
plot (\emph{i}), \(x_{i,d}\).
\new{Then we standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_G x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.}
We then found the difference between the standardized ambient soil
moisture and the standardized drought and irrigation soil moisture
within each plot group. These transformations ensured that the treatment
effects in each plot were appropriately scaled by the local ambient
conditions within each plot group.

We then modeled the daily deviation from ambient conditions of the
drought and irrigation treatments using a linear mixed effects model
with independent variables for treatment, season (winter, spring,
summer, fall), rainfall, and all two-way interactions. Rainy days were
defined as any day in which precipitation was recorded and average
temperature was above 3\(^{\circ}\)C. The day immediately following
rainfall was also classified as rainy. We fit the model using the
\texttt{lme4::lmer()} function \citep{Bates2015} in R \citep{R2016},
with random effects for plot group and date. We weighted observations by
the number of unique sensors or spot measurements that were taken in
each plot on that day. We then used the model to predict the average
daily soil moisture in the treated plots based on the average daily
ambient soil moisture. We could only predict soil moisture in the
treated plots on days for which we took at least one ambient soil
moisture measurement. See \citet{Kleinhesselink2017b} for complete
details on our approach to estimating soil moisture.

>>>>

\new{We first averaged the observed soil moisture for each day ($d$) and plot ($i$), $x_{i,d}$.
Plots were assigned to plot groups consisting of a control, a drought, and an irrigation plot.
We standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_n x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.
These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.
We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group: $\Delta x^*_{\text{trt.}} = x^*_{\text{control}} - x^*_{\text{trt.}}$.}

We then modeled the daily deviation \new{($\Delta x^*_\text{trt.}$)}
from ambient conditions of the drought and irrigation treatments using a
linear mixed effects model with independent variables for treatment,
season (winter, spring, summer, fall), rainfall, and all two-way
interactions. Rainy days were defined as any day in which precipitation
was recorded and average temperature was above 3\(^{\circ}\)C. The day
immediately following rainfall was also classified as rainy. We fit the
model using the \texttt{lme4::lmer()} function \citep{Bates2015} in R
\citep{R2016}, with random effects for plot group and date. We weighted
observations by the number of unique sensors or spot measurements that
were taken in each plot on that day. We then estimated average daily
soil moisture in the treated plots based on the average daily ambient
soil moisture
\new{and the model predicted daily deviation for each treatment ($\Delta x^*_\text{trt.}$)}.
We could only predict soil moisture in the treated plots on days for
which we took at least one ambient soil moisture measurement. See
\citet{Kleinhesselink2017b} for complete details on our approach to
estimating soil moisture.
